---
title: "Predictors of Weight in RomanianInfants: Weight-for-age z-scores (WAZ)"
author: "Adrian Rus"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Setting Up the Environment: R Packages Installation----------------------
# install.packages("tidyverse")
# install.packages("moments")
# install.packages(c("lmtest", "car", "carData"))

library(tidyverse)
library (moments)
library(lmtest)
library(car)
library(carData)

```

## 1. Research Hypotheses/Questions/Objectives

The objective of the present study was to explore the relationship between infants' weight and various demographics (i.e., infants' mother location, marital status, social economic status, and age; and infants being born on term, age, anemia status, rank into the family and dietary patterns).

Identifying the connection between such factors could help practitioners in developing public health policies and interventions in Romania.

## 2. Data Exploration

### 2.1. Setting the Working Directory

Check the working directory

```{r}
getwd()
```

For setting a new working directory use the \`setwd()1

setwd("C:\\\\Users\\\\adria\\\\Documents\\\\waz_infants")

List Files in Directory To see a list of all files in the current working directory, you can use

```{r}
list.files()
```
