---
title: "Predictors of Weight in Romanian Infants: Weight-for-age z-scores (WAZ)"
author: "Adrian Rus"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Setting Up the Environment: R Packages Installation----------------------
# install.packages("tidyverse")
# install.packages("moments")
# install.packages(c("lmtest", "car", "carData"))

library(tidyverse)
library (moments)
library(lmtest)
library(car)
library(carData)

```

## 1. Research Hypotheses/Questions/Objectives

The objective of the present study was to explore the relationship between infants' weight and various demographics (i.e., infants' mother location, marital status, social economic status, and age; and infants being born on term, age, anemia status, rank into the family and dietary patterns).

Identifying the connection between such factors could help practitioners in developing public health policies and interventions in Romania.

## 2. Data Exploration

### 2.1. Setting the Working Directory

Check the working directory

```{r}
getwd()
```

### 2.2. For setting a new working directory use the `setwd()`

```{r}

setwd("C:\\Users\\adria\\Documents\\waz_infants\\data")

```

List Files in Directory To see a list of all files in the current working directory, you can use

```{r}
list.files()
```

### 2.3. Import and check the dataset

```{r}

library(readr)
waz_final <- read_csv("C:/Users/adria/Documents/waz_infants/data/waz_final.csv")
View(waz_final)

```

#### 2.3.1. Basic inspection of dataset

```{r}
# View the first 6 rows
head(waz_final)
```

```{r}
# View the last 6 rows
tail(waz_final)
```

```{r}
# View the structure of dataset
str(waz_final)

```

```{r}
# Summarize the dataset
summary(waz_final)
```

#### 2.3.2. Check the tidiness of the dataset

```{r}
# Checking the missing values. 
# Tidy data should have a consistent structure, but it can contain missing values.
sum(is.na(waz_final))
```

```{r}
# Tidy data should ideally not have duplicated rows unless the repetition is meaningful.        
sum(duplicated(waz_final))
```

```{r}
# Column names should be clear and descriptive but not too lengthy.
names(waz_final)
```

#### 2.3.3. Check the data types and dimensionality

```{r}
# Tells you whether the object is a data frame, matrix, vector, list, etc.
class(waz_final)
```

```{r}
# Returns the class (type) of each column in the data frame.
sapply(waz_final, class)
```

```{r}
# Returns the number of columns.
ncol(waz_final)
```

```{r}
# Returns the number of rows and columns in a data frame.
dim(waz_final)

```

```{r}
# Select only numerical columns

numerical_columns <- waz_final[, sapply(waz_final, is.numeric)]

head(numerical_columns) # Show the first few rows of the numerical columns

View(numerical_columns) # It shows all the variables
```

```{r}
# Select only categorical columns (factor or character)
categorical_columns <- waz_final[, sapply(waz_final, function(col) is.factor(col) || is.character(col))]
# Show the first few rows of the categorical columns
head(categorical_columns)
```

#### 2.3.4. Specific variable content

```{r}
# Shows unique values in a specific column.

unique(waz_final$mAge2c)
```

```{r}
# Allows to identify the unique data in all columns

unique_values_all_columns <- lapply(waz_final, unique)

str(unique_values_all_columns)
```

## 3. Dependent Variable (DV) and Data Processing

***Dependent variable (DV)***: cWageZ (weight-for-age z-scores; WAZ)

***Independent variable (IV)***:

### 3.1. Visual check of the DV

```{r}
# Histogram for DV
library(ggplot2)
ggplot(waz_final, aes(x = cWageZ)) + 
  geom_histogram(bins = 50, col= "white") +
  ggtitle("Histogram")
```

```{r}
# Histogram adding the normal distribution curve
# Calculate mean and standard deviation of the data
data_mean <- mean(waz_final$cWageZ, na.rm = TRUE)
data_sd <- sd(waz_final$cWageZ, na.rm = TRUE)
# Create the ggplot
ggplot(waz_final, aes(x = cWageZ)) +
  geom_histogram(aes(y = after_stat(density)), bins = 50, col= "white") + 
  stat_function(fun = dnorm, args = list(mean = data_mean, sd = data_sd), color = "red") +
  ggtitle("Histogram with Normal Distribution Curve")
```

```{r}
# Boxplot to identify outliers and understand the data's spread.
ggplot(waz_final, aes(y = cWageZ)) + geom_boxplot()
ggplot(waz_final, aes(y = cWageZ)) + 
  geom_boxplot(outlier.colour = "red")

```

```{r}
# Density Plot: To visualize the distribution.
ggplot(waz_final, aes(x = cWageZ)) + geom_density()

```

```{r}
# Q-Q Plot: To assess normality.
ggplot(waz_final, aes(sample = cWageZ)) + stat_qq()
```

```{r}
#Pair Plot: To visualize relationships between multiple numerical variables, if applicable.

# Filter only numeric columns
# waz_final_numeric <- waz_final[, sapply(waz_final, is.numeric)]


# Create the pairs plot
# pairs(waz_final_numeric)
```

### 3.2. Compute central tendency, dispersion, skewness, kurtosis, and normality (Shapiro-Wilk P-value)

#### To be continued.